---
title: 'Peer Graded Assignment: Week 4 Machine Learning'
author: "Nicole Lakens"
date: "4 oktober 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Description of the project

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.

## What I should submit

The goal of my project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. We can use any of the other variables to predict with. I should create a report describing how I built the model, how I used cross validation, what I think the expected out of sample error is, and why I made the choices I did. I will also use my prediction model to predict 20 different test cases.

# Building the model

## Used packages

First, I make sure the packages that I will be using are active.


```{r packages}
library(RANN)
library(dplyr)
library(caret)
```

## Reading in the data and check how it looks

I did this twice because I found out that some variables had text in it like "#DIV/0". So I added the "na.strings"-syntax to the read.csv.

```{r reading_checking_data}
training <- read.csv("C:/Users/nlake/Downloads/pml-training.csv", na.strings = c("NA", "", "#DIV/0!"))
testing <- read.csv("C:/Users/nlake/Downloads/pml-testing.csv", na.strings = c("NA", "", "#DIV/0!"))

summary(training)
head(training)
table(training$classe)

```

## Deleting variables

Column X is a row indicator; let's delete this one. User name cannot be a predictor in this model; let's delete the user name.
A timestamp does not look like a variable that can predict a movement; let's delete the variables that we will not use. 

```{r delete_vars}
training$X <- NULL
testing$X <- NULL

training$user_name <- NULL
testing$user_name <- NULL

training$raw_timestamp_part_1 <- NULL
testing$raw_timestamp_part_1 <- NULL


training$raw_timestamp_part_2 <- NULL
testing$raw_timestamp_part_2 <- NULL


training$cvtd_timestamp <- NULL
testing$cvtd_timestamp <- NULL


training$new_window <- NULL
testing$new_window <- NULL


training$num_window <- NULL
testing$num_window <- NULL
```

## Splitting the trainingset

Split the trainingset in a trainingset and validationset, so that we can calculate the out of sample errorrate.

```{r split}
set.seed(800)
partit <- createDataPartition(y=training$classe, p=0.75, list=FALSE)
training_part <- training[partit, ]
validation_part <- training[-partit, ]
```

## Check the distribution of the var to be predicted

Create a table and a plot so we can see how the classe-variable is distributed.

```{r plot}
table(training_part$classe)
table(training_part$classe)/nrow(training_part)
plot(training_part$classe)
```

## Imputation and (near) zero variance variables

Check if NA's are present in the dataset.

```{r nas}
table(is.na(training_part))
```

Yes, there are NA's, so we have to impute the NA's. I am using the preprocess function from Caret to impute the NA's. I will also remove zero variance and near zero variance columns.
```{r nzv}
nzv <- nearZeroVar(training_part)
nzv
training_part <- training_part[,-nzv]
validation_part <- validation_part[,-nzv]
testing_part <- testing[,-nzv]

preproc <- preProcess(training_part, method = c('medianImpute'))
preproc
```

Recreate dataset with imputed values. Same with the validation-data and the testing-data.

```{r predict_preproc}
training_pp<- predict(preproc, training_part)
validation_pp<- predict(preproc, validation_part)
testing_pp<- predict(preproc, testing_part)
```

Also, set the classe-variable as factor.

```{r fact}
training_pp$classe <- as.factor(training_pp$classe)
validation_pp$classe <- as.factor(validation_pp$classe)
testing_pp$Classe <- c('')
```

## The model

The k-fold cross validation method involves splitting the dataset into k-subsets. For each subset is held
out while the model is trained on all other subsets. This process is completed until accuracy is determine
for each instance in the dataset, and an overall accuracy estimate is provided. It is a robust method for
estimating accuracy, and the size of k and tune the amount of bias in the estimate,
with popular values set to 3, 5, 7 and 10.

I will be building a random forest using Caret; I changed the default number of trees from 500 to 100 because of the limited capacity of my laptop. For the same reason, I used 3 folds.

```{r model}
train_control <- trainControl(method="cv", number=3)

model <- train(classe~., data=training_pp, method="rf", trainControl=train_control, ntree = 100)

model

pred <- predict(model)

varImp(model)
```

## Confusion matrix for training set

To check how the model performs on the training set, we will create a confusion matrix.

```{r cf_training}
cf <- caret::confusionMatrix(pred,training_pp$classe)
cf
```

```{r training_errorrate}
training_errorrate<- 1-cf$overall[[1]]
```

The error rate for the training set is `r training_errorrate`.

## Apply to validation set

Now we use our model to predict the validation set; in this way, we can see what our expected out of sample error rate. We will create the confusionmatrix to do so.

```{r predict_val}

val_pred <- predict(model, newdata = validation_pp)
cf_val<- caret::confusionMatrix(val_pred, validation_pp$classe)
cf_val
```
```{r val_errorrate}
val_errorrate <- 1-cf_val$overall[[1]]
val_errorrate
```

The out of sample error rate is `r val_errorrate`. So this is the errorrate we can expect to see when predicting to another unseen dataset.

## Apply to testing set

The last part of the project is to predict 20 testcases.

```{r pred_test}
testing_pred <- predict(model, newdata = testing_pp)
testing_pred
```


# Thank you for reviewing my project :-) 










